```{bash}
awk 'BEGIN {OFS=FS="\t"} {if(FNR==43||FNR==44||FNR==52||FNR==53||FNR==54||FNR==67) print $0}' GPL11154.txt|cut -f 1 --complement> sample1.txt
awk 'BEGIN {OFS=FS="\t"} {if(FNR==43||FNR==44||FNR==52||FNR==53||FNR==54||FNR==67) print $0}' GPL18573.txt|cut -f 1 --complement> sample2.txt
paste sample1.txt sample2.txt >sample.txt
```
```{r}
samplefile=c("sample.txt","sample1.txt","sample2.txt")
sample_format<-function(samplefile){
sample <- read.table(samplefile, quote="\"", sep="\t",stringsAsFactors = F)
sample<-rbind(sample,rep("",ncol(sample)))
rownames(sample)<-c("samplenames","sampleid","age","sex","oa grade","platform","conditions")
sample<-data.frame(t(sample),stringsAsFactors = F)
sample$age<-sub("age: ","", sample$age)
sample$sex<-sub("Sex: ","", sample$sex)
sample$oa.grade<-sub("oa grade: ","",sample$oa.grade)
sample$conditions[grep("^Norm|^norm",sample$samplenames)]<-"Norm"
sample$conditions[grep("^OA",sample$samplenames)]<-"OA"
sample$studyid<-"GSE114007"
write.table(sample,samplefile,sep="\t",quote=F,col.names=T,row.names = F)
}
for(i in samplefile) sample_format(i)
```
```{r}
library(DESeq2)
library("RColorBrewer")
library("gplots")
library("amap")
library("ggplot2")
library("BiocParallel")
library("ggrepel")
```
```{r}
sample <- read.table("GSE114007_sample.txt", quote="", sep="\t",header = T)
sample<-sample[order(sample$conditions),]
sample_rowname <- sample$samplenames
sample<-sample[,-1]
rownames(sample)<-sample_rowname
```
```{r}
#GSEA group
cat(c(nrow(sample),"2 1\n"),file="GSEA_group.cls" )
cat(c("#",levels(as.factor(sample$conditions)),"\n"),file="GSEA_group.cls",append = TRUE)
cat(as.character(sample$conditions),file="GSEA_group.cls",append = TRUE)
#WGCNA
sample_wgcna<-sample
sample_wgcna$OA<-ifelse(sample_wgcna$conditions=="OA",1,0)
sample_wgcna$Norm<-ifelse(sample_wgcna$conditions=="Norm",1,0)
sample_wgcna$sex<-ifelse(sample_wgcna$sex=="F",0,1)
sample_wgcna<-sample_wgcna[,c("sex","age","OA","Norm")]
write.table(sample_wgcna,"wgcna_trait.txt",sep = "\t",quote = F)
```
```{r}
data1 <- read.csv("GSE114007_raw_counts_Norm.csv", header=T, row.names=1, com='', quote='', check.names=F, sep=",")
data2 <- read.csv("GSE114007_raw_counts_OA.csv", header=T, row.names=1, com='', quote='', check.names=F, sep=",")
data<-data.frame(data1,data2)
rm(data1,data2)
data<-data[,colnames(data)%in%rownames(sample)]
data<-data[,match(rownames(sample),colnames(data))]
data_output<-data.frame(symbol=rownames(data),data,stringsAsFactors = F)
```
```{r}
#Unify the gene symbols
write.table(rownames(data),"to_match.txt",sep="/t",col.names = F,row.names = F,quote = F)
hsc<-read.csv("hgnc-symbol-check.csv",header=T,sep=",",skip=1,stringsAsFactors = F)
hsc<-hsc[!duplicated(hsc$Input),]
hsc_ps<-hsc[hsc$Match.type=="Previous symbol"|hsc$Match.type=="Alias symbol",]
hsc_w<-hsc[hsc$Match.type=="Entry withdrawn",]
for(i in hsc_ps$Input) data_output$symbol[which(data_output$symbol==i)]<-hsc_ps$Approved.symbol[which(hsc_ps$Input==i)]
for(i in hsc_w$Input) data_output$symbol[which(data_output$symbol==i)]<-""
data_output<-data_output[which(data_output$symbol!=""),]
data_output<-aggregate(data_output[,-1],by=list(symbol=data_output$symbol),FUN=median)
write.table(data_output,"GSE114007_match.txt",col.names = T,row.names = F,sep = "\t",quote = F)
data<-read.table("GSE114007_match.txt",header = T,row.names = 1,sep = "\t",quote ="")
data <- data[rowSums(data)>nrow(sample)/2,]
data<-round(data)
```
##Build DESeq2 object
```{r}
ddsFullCountTable <- DESeqDataSetFromMatrix(countData = data,colData = sample,  design = ~ platform + conditions)
dds <- DESeq(ddsFullCountTable)
```
##Normalize the data and order the genes according to mad value
```{r}
normalized_counts <- counts(dds, normalized=TRUE)
normalized_counts_mad <- apply(normalized_counts, 1, mad)
normalized_counts <- normalized_counts[order(normalized_counts_mad, decreasing=T), ]
normalized_counts_output<- data.frame(symbol=rownames(normalized_counts), normalized_counts)
write.table(normalized_counts_output, file="DESeq2.normalized.txt",quote=F, sep="\t", row.names=F, col.names=T)  
rld <- rlog(dds, blind = F)
rlogMat <- assay(rld)
rlogMat<-limma::removeBatchEffect(rlogMat,sample$platform)
rlogMat <- rlogMat[order(normalized_counts_mad, decreasing=T), ]
rlogMat_output<-data.frame(symbol=rownames(rlogMat),rlogMat)
write.table(rlogMat_output, file="DESeq2.normalized.rlog_batch_removed.txt",quote=F, sep="\t", row.names=F, col.names=T)
```
## Expression profile of normalized samples
```{r}
rlog_mat_melt <- reshape2::melt(rlogMat_output, id.vars = c('symbol'))
ggplot(rlog_mat_melt, aes(x=variable, y=value)) + 
  geom_boxplot(aes(color=variable)) +
  geom_violin(aes(fill=variable), alpha=0.5) + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.title.x = element_blank()) + ylab("rLog transformed expression value")
```
##Hierarchical clustering to determine the similarity of samples and differences within groups
```{r}
hc <- amap::hcluster(t(rlogMat), method="pearson")
plot(hc,labels=sample$conditions)
```
##Calculate pearson correlation values and draw the heatmap
```{r,fig.height=15,fig.width=13.5}
pearson_cor <- as.matrix(cor(rlogMat, method="pearson"))
pearson_cor <- pearson_cor[hc$order, hc$order]
pearson_cor_output = data.frame(sample=rownames(pearson_cor), pearson_cor)
# Melt the correlation matrix
melted_cormat <- reshape2::melt(pearson_cor, na.rm = TRUE)
col = colorRampPalette(colors = c("blue","white","red"))(100)
# Create a ggheatmap
p <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  #labs(title = "The pearson correlation of each sample")+
  scale_fill_gradientn(colours=col, name="Pearson correlation") +
  theme_classic() +
  coord_fixed() + 
  theme(
  axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  legend.justification = c(1, 1),
  legend.position = "right",
  legend.direction = "vertical")
  guides(fill = guide_colorbar(barwidth = 1, barheight = 8,
                title.position = "top"))
p
```
##PCA analysis of samples
```{r}
rlogMat<-read.table("DESeq2.normalized.rlog.txt",sep="\t",row.names = 1,header = T)
formulaV <- c("conditions","platform")
pca_mat<-rlogMat
pca_mat <- as.data.frame(t(pca_mat))
pca <- prcomp(pca_mat, scale=T)
pca_x = pca$x
pca_individual = data.frame(samp=rownames(pca_x), pca_x, sample)
pca_percentvar <- formatC(pca$sdev^2 * 100 / sum( pca$sdev^2))
pca_percentvar <- data.frame(PC=colnames(pca_x), Variance=pca_percentvar)
color<-c(Norm="blue",OA="red")
if (length(formulaV)==1) {
  p <- ggplot(pca_individual, 
              aes(PC1, PC2, color=conditions))
} else if (length(formulaV)==2) {
  p <- ggplot(pca_individual, 
              aes(PC1, PC2, color=conditions,
                  shape=platform))
}

p <- p + geom_point(alpha=1) +
  guides(color=guide_legend(title="Condition",order = 1),
         shape=guide_legend(title = "Platform",order = 0))+
  #labs(title = "PCA plot")+
	xlab(paste0("PC1: ", pca_percentvar[1], "% variance")) +
	ylab(paste0("PC2: ", pca_percentvar[2], "% variance")) +
	theme_classic() +
	theme(legend.position="right")+
  scale_color_manual(values = color)+
  ggrepel::geom_text_repel(aes(label=samp), show.legend=F) 
p
```
```{r}
sampleA = "OA"
sampleB = "Norm"
contrastV <- c("conditions", sampleA, sampleB)
res <- results(dds,  contrast=contrastV)
resultsNames(dds)
contrastV <- paste("conditions", sampleA, "vs",sampleB,sep = "_")
res <- results(dds, name = contrastV)
res
```
##Get the mean of the first set of data
```{r}
baseA <- counts(dds, normalized=TRUE)[, colData(dds)$conditions == sampleA]

if (is.vector(baseA)){
    baseMeanA <- as.data.frame(baseA)
} else {
    baseMeanA <- as.data.frame(rowMeans(baseA))
}
colnames(baseMeanA) <- sampleA
head(baseMeanA)
```
##Get the mean of the second set of data
```{r}
baseB <- counts(dds, normalized=TRUE)[, colData(dds)$conditions == sampleB]
if (is.vector(baseB)){
        baseMeanB <- as.data.frame(baseB)
} else {
        baseMeanB <- as.data.frame(rowMeans(baseB))
}
colnames(baseMeanB) <- sampleB
head(baseMeanB)
```
```{r}
res <- cbind(baseMeanA, baseMeanB, as.data.frame(res))
head(res)
```
```{r}
res <- cbind(symbol=rownames(res), as.data.frame(res))
#res$baseMean <- rowMeans(cbind(baseA, baseB))
# The value of NA after correction is 1
res$padj[is.na(res$padj)] <- 1
# Sorted by pvalues, put genes with big difference in front
res <- res[order(res$pvalue),]
head(res)
```
```{r}
# generate file name
comp<-paste("DESeq2",sampleA,"vs",sampleB,sep = ".")
file_base <- paste(comp, "results.txt",sep=".")
res_output <- as.data.frame(subset(res,select=c('symbol',sampleA,sampleB,"baseMean",'log2FoldChange','pvalue', 'padj')))
write.table(res_output, file=file_base, sep="\t", quote=F, row.names=F)

# Differentially expressed gene（DEG） screening，padj<0.05
res_de <- subset(res, res$padj<0.05, select=c('symbol', sampleA,sampleB, 'log2FoldChange', 'padj'))

# Foldchange >= 1
res_de_up <- subset(res_de, res_de$log2FoldChange>= 1)
comp1<- paste("DESeq2",sampleA,"higher_than",sampleB, sep=".")
res_de_up_symbol <- subset(res_de_up, select=c("symbol"))
res_de_dw <- subset(res_de, res_de$log2FoldChange<=(-1)*1)
comp2 <- paste("DESeq2",sampleA,"lower_than",sampleB, sep=".")
res_de_dw_symbol <- subset(res_de_dw, select=c("symbol"))
res_de<-rbind(res_de_up,res_de_dw)
write.table(res_de, 
            file=paste(comp,'DE.txt', sep="."), sep="\t", quote=F, row.names=F)
if(dim(res_de_up_symbol)[1]>0) 
  res_de_up_symbol_l <- cbind(res_de_up_symbol, 
                              type=paste(sampleA,"higher_than",sampleB, sep="."))
if(dim(res_de_dw_symbol)[1]>0) 
  res_de_dw_symbol_l <- cbind(res_de_dw_symbol, 
                              type=paste(sampleA,"lower_than",sampleB, sep="."))
res_de_symbol_l<-rbind(res_de_up_symbol_l,res_de_dw_symbol_l)
write.table(res_de_symbol_l, 
              file=paste(comp,'DE.symbol.txt', sep="."),sep="\t",quote=F, row.names=F, col.names=F, append=F)
```
```{r}
res_output<-read.table("DESeq2.OA.vs.Norm.results.txt",sep = "\t",quote = "",header = T)
res_output$level <- ifelse(res_output$padj<=0.05, 
                           ifelse(res_output$log2FoldChange>=1, 
                                  paste(sampleA,"Up"), 
                                  ifelse(res_output$log2FoldChange<=1*(-1),
                                         paste(sampleB,"Up"), "NoDiff")) , "NoDiff")
res_output$level<-factor(res_output$level,levels=c(paste(sampleB,"Up"),"NoDiff",paste(sampleA,"Up")),labels = c("Down","NoDiff","Up"))
res_output_sig<-res_output[res_output$padj<0.05,]
OA_top<-head(res_output_sig$symbol[order(res_output_sig$log2FoldChange,decreasing = T)],10)
OA_top<-as.character(OA_top)
Norm_top<-head(res_output_sig$symbol[order(res_output_sig$log2FoldChange,decreasing = F)],10)
Norm_top<-as.character(Norm_top)
top<-c(OA_top,Norm_top)
res_output$label<-""
res_output$label[match(top,res_output$symbol)]<-top
res_output$log_padj <- (-1) * log10(res_output$padj)
color<-c("red","blue","grey")
names(color)<-c("Up","Down", "NoDiff") 
p <-ggplot(res_output,aes(x=log2FoldChange,y=log_padj))+
  geom_point(aes(color=level,size=baseMean),alpha=0.5) +
  scale_color_manual(values =color,
                     limits=c("Down","NoDiff","Up"))+
  scale_size(guide = F)+
  theme_classic() +
  geom_hline(yintercept = -log10(0.05), lty=4,col="grey",lwd=0.6) +
  geom_vline(xintercept = c(-1, 1), lty=4,col="grey",lwd=0.6)+
  xlab(expression(paste("Log"[2],"(Fold-change)")) ) + ylab(expression(paste("- Log"[10],"(FDR)"))) +#labs(title = "Volcano plot of differentially expressed genes")+
  guides(color=guide_legend(label.position="top",nrow = 1)) +
  theme(legend.position = "top",legend.title = element_blank())+
  geom_text_repel(aes(label=label))
```
## Sort genes by log2FolcChange
```{r}
res_output_line <- res_output[order(res_output$log2FoldChange),]
res_output_line$x <- 1:nrow(res_output_line)
res_output_line[abs(res_output_line$log2FoldChange)<5 | res_output_line$level=="NoDiff", "symbol"] <- NA
p <- ggplot(res_output_line, aes(x=x, y=log2FoldChange)) + 
  geom_point(aes(color=log2FoldChange)) + 
  scale_color_gradient2(low="blue", mid="white", high= "red", midpoint = 0) + 
  theme_classic() + geom_hline(yintercept = 0, linetype="dotted")
p +  geom_text_repel(aes(label=symbol))
```
## Top20 significantly different genes
```{r}
res_de_up<-read.table("DESeq2.OA.higher_than.Norm.DE.txt",sep = "\t",quote = "",header = T)
res_de_dw<-read.table("DESeq2.OA.lower_than.Norm.DE.txt",sep = "\t",quote = "",header = T)
res_de_up_top20_id <- as.vector(head(res_de_up$symbol,10))
res_de_dw_top20_id <- as.vector(head(res_de_dw$symbol,10))
res_de_top20 <- c(res_de_up_top20_id, res_de_dw_top20_id)
res_de_top20
```
##DEG heatmap
```{r}
normalized_counts<-read.table("DESeq2.normalized.txt",sep = "\t",quote = "",header = T,row.names = 1)
res_de_top20_expr <- normalized_counts[res_de_top20,]
res_de_top20_expr<-as.matrix(res_de_top20_expr)
sample <- read.table("GSE114007_sample.txt", quote="", sep="\t",header = T)
sample<-sample[order(sample$conditions),]
sample_rowname <- sample$samplenames
sample<-sample[,-1]
rownames(sample)<-sample_rowname
ann_col<-sample[,c("conditions","sex","age")]
colnames(ann_col)<-c("Condition","Sex","Age")
library(pheatmap)
ann_colors = list(
    conditions = c(Norm="blue", OA="red"))
p<-pheatmap(res_de_top20_expr, 
         color=col,
         breaks = seq(-2,2,length.out = 100),
         cluster_row=F, 
         cluster_cols = F ,
         show_colnames = F,
         scale ="row",
         annotation_col =ann_col,
         #main ="Heatmap for differentially expressed genes between normal and OA tissues ordered by adjusted p values"
         )
p
```
##DEG scatter plot
```{r}
res_de_top20_expr2 <- data.frame(Gene=rownames(res_de_top20_expr), res_de_top20_expr)
head(res_de_top20_expr2)
```
```{r}
library(reshape2)
res_de_top20_expr2 <- melt(res_de_top20_expr, id=c("Gene"))
colnames(res_de_top20_expr2) <- c("Gene", "Sample", "Expression")
head(res_de_top20_expr2)
```
```{r}
ggplot(res_de_top20_expr2, aes(x=Gene, y=Expression)) +
  geom_point(aes(color=Sample), alpha=0.5) + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.title.x = element_blank()) + 
  ylab("Normalized xpression value") + 
  scale_y_log10()
```
##DEG scatter plot
```{r}
# Add group information attribute
res_de_top20_expr2$Group <- sample[match(res_de_top20_expr2$Sample, rownames(sample)),"conditions"]
head(res_de_top20_expr2)
color<-c("red","blue")
names(color)<-c(sampleA,sampleB)
ggplot(res_de_top20_expr2, aes(x=Gene, y=Expression)) + 
  geom_point(aes(color=Group), alpha=0.5) +
  scale_color_manual(values = color)+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.title.x = element_blank()) + ylab("Normalized xpression value") + scale_y_log10()
```
